Nanophotonic Inverse Design

### 2011-12-13 presentations/inv_dev_plan_dec-2011/paper.tex
Working on apendix still.

\subsubsection{Linear $f(x)$}
We investigate the choice $f(x) = c^T x$, 
    where $c \in \reals^{n \times 1}$
    and $x \in \reals^{n \times 1}$.
Here,
    \begin{subequations}\begin{align}
    \nabla f(x) &= c \\
    \nabla^2 f(x) &= 0
    \end{align}\end{subequations}
    which results in
    \begin{subequations}\begin{align}
    \nabla_x \mathcal{L} &= c + \rho A^T (A x - b + \inv{\rho} y) 
        = \rho A^T (Ax - b + \inv{\rho}(y + \tilde{c})) \\
    \nabla_{xx}^2 \mathcal{L} &= \rho A^T A,
    \end{align}\end{subequations}
    where $\tilde{c} = A^{-T} c$.
A single Newton step is
    \begin{equation}\begin{split}
    \Delta x &= (\nabla_{xx}^2 \mathcal{L})^{-1} \nabla_x \mathcal{L} \\
        &= (\rho^{-1} A^{-1} A^{-T}) 
            A^T (A x - b + \inv{\rho}(y + \tilde c)) \\
        &= A^{-1}(A x - b + \inv{\rho}(y + \tilde c)).
    \end{split}\end{equation}

@@

Newton's method for a linearly-constrained quadratic function is
    \begin{equation}
    \begin{bmatrix} \nabla_{xx}^2 \mathcal{L} & C_2 \\ C_2^T & 0 \end{bmatrix} 
    \begin{bmatrix} \Delta x \\ v \end{bmatrix} = 
    \begin{bmatrix} -\nabla_x\mathcal{L} \\ C_2 x - d_2 \end{bmatrix}.
    \end{equation}
We assume that 
    \begin{subequations}\begin{align}
    (\nabla_{xx}^2 \mathcal{L})^{-1} &= A^{-1} M A^{-T} \\
    \nabla_x \mathcal{L} &= A^T z
    \end{align}\end{subequations}
    and solve for $\Delta x$ 
    using block elimination (see appendix \ref{block elim}),
    \begin{subequations}\begin{align}
    v &= -(\tC_2^T M \tC_2)^{-1}(\tC_2^T Ax - d_2 + \tC_2^T Mz) \\
    \Delta x &= -A^{-1} M (z + \tC_2 v),
    \end{align}\end{subequations}
    where $\tC_2 = A^{-T} C_2$.
    If we condense the block elimination procedure to a single step,
    we obtain
    \begin{equation}
    \Delta x = -A^{-1} M (z - 
        \tC_2 (\tC_2^T M \tC_2)^{-1}(\tC_2^T Ax - d_2 + \tC_2^T Mz)).
    \end{equation}

@@

To deal with the concavity of $f(x)$ we would like to use
    a truncated Newton method,
    \begin{equation}
    \hat{\mathcal{L}}(x,p,y) = \mathcal{L}(x,p,y) + \frac{\eta}{2} \|x - x_0\|^2
    \end{equation}
    or even
    \begin{equation}
    \Delta x = -(\nabla_{xx}^2 \mathcal{L} + \eta I)^{-1}
        \nabla_x \mathcal{L},
    \end{equation}
    where the coefficient $\eta$ is increased in order to 
    ensure that the augmented Hessian is positive definite 
    (so that $(\nabla_x\mathcal{L})^T\Delta x > 0$).
However, this generally makes inverting the Hessian intractable.

@@

\subsubsection{Linearly constrained $f(x)$ with arbitrary phase}
We now investigate another special case of $f(x)$, 
    namely, when $f(x)$ implements the equality constraint 
    $C_2^T x = e^{i \phi}d_2$,
    where $C_2 \in \reals^{n \times p_2}$ and $p_2 \ll n$.
This situation corresponds to
    the addition of an arbitrary phase $e^{i\phi}$
    to a linear equality constraint.

We simply note that this case can be handled by computing
    \begin{equation}
    \Delta x = -A^{-1} M (z - 
        \tC_2 (\tC_2^T M \tC_2)^{-1}(\tC_2^T Ax - e^{i\phi}d_2 + \tC_2^T Mz)).
    \end{equation}
    for different choices of $\phi$.
The total computational cost is then simply $n_\phi + p_2$ solves,
    where $n_\phi$ is the number of values of $\phi$ which are tried.

\subsection{Multi-mode ADMM}
Since ADMM explicitly separates updating the field and structure variables,
    handling multiple fields is very naturally included in the formulation.
Namely, multiple fields ($x_i$) are updated simultaneously, in parallel,
    and then care ``connected'' by updating a single $p$ 
    which takes all the $x_i$ into account.

\begin{appendix}
\section{Constructing the relevant matrices and vectors}
\section{Solving the matrices}
\section{Linear algebra definitions}
\subsection{Gradient and Hessian calculations}
\subsection{Matrix inversion lemma}\label{inv lemma}
\begin{equation}
(A + UCV)^{-1} = A^{-1} - A^{-1}U (C^{-1} + V A^{-1} U)^{-1} V A^{-1}
\end{equation}
\subsection{Block elimination of a matrix}\label{block elim}
\begin{equation}
\begin{bmatrix} A_{11} & A_{12} \\ A_{21} & A_{22} \end{bmatrix} 
\begin{bmatrix} x_1 \\ x_2 \end{bmatrix} = 
\begin{bmatrix} b_1 \\ b_2 \end{bmatrix}
\end{equation}
\begin{subequations}\begin{align}
A_{11} x_1 + A_{12} x_2 &= b_1 \\
x_1 &= A_{11}^{-1} (b_1 - A_{12} x_2)
\end{align}\end{subequations}
\begin{subequations}\begin{align}
A_{21} x_1 + A_{22} x_2 &= b_2 \\
A_{21} A_{11}^{-1} (b_1 - A_{12} x_2) + A_{22} x_2 &= b_2 \\
(A_{22} - A_{21} A_{11}^{-1} A_{12}) x_2 &= b_2 - A_{21} A_{11}^{-1} b_1
\end{align}\end{subequations}
Typically, one computes $A_{11}^{-1} A_{12}$ and $A_{11}^{-1} b_1$, 
    and then solves for the rest of the system.
\end{appendix}
\begin{thebibliography}{99}
\bibitem{ADMM} Boyd group ADMM paper
\end{thebibliography}
\end{document}

>>>>

\subsubsection{Linear $f(x)$}
We investigate the choice $f(x) = c^T x$, 
    where $c \in \reals^{n \times 1}$
    and $x \in \reals^{n \times 1}$.
Here,
    \begin{subequations}\begin{align}
    \nabla f(x) &= c \\
    \nabla^2 f(x) &= 0
    \end{align}\end{subequations}
    which results in
    \begin{subequations}\begin{align}
    \nabla_x \mathcal{L} &= c + \rho A^T (A x - b + \inv{\rho} y) 
        = \rho A^T (Ax - b + \inv{\rho}(y + \tilde{c})) \\
    \nabla_{xx}^2 \mathcal{L} &= \rho A^T A,
    \end{align}\end{subequations}
    where $\tilde{c} = A^{-T} c$.
A single Newton step is
    \begin{equation}\begin{split}
    \Delta x &= (\nabla_{xx}^2 \mathcal{L})^{-1} \nabla_x \mathcal{L} \\
        &= (\rho^{-1} A^{-1} A^{-T}) 
            \rho A^T (A x - b + \inv{\rho}(y + \tilde c)) \\
        &= A^{-1}(A x - b + \inv{\rho}(y + \tilde c)).
    \end{split}\end{equation}

@@

Newton's method for a linearly-constrained quadratic function is
    \begin{equation}
    \begin{bmatrix} \nabla_{xx}^2 \mathcal{L} & C_2 \\ C_2^T & 0 \end{bmatrix} 
    \begin{bmatrix} \Delta x \\ v \end{bmatrix} = 
    \begin{bmatrix} -\nabla_x\mathcal{L} \\ C_2^T x - d_2 \end{bmatrix}.
    \end{equation}
We assume that 
    \begin{subequations}\begin{align}
    (\nabla_{xx}^2 \mathcal{L})^{-1} &= A^{-1} M A^{-T} \\
    \nabla_x \mathcal{L} &= A^T z
    \end{align}\end{subequations}
    and solve for $\Delta x$ 
    using block elimination (see appendix \ref{block elim}),
    \begin{subequations}\begin{align}
    v &= -(\tC_2^T M \tC_2)^{-1}(\tC_2^T Ax - d_2 + \tC_2^T Mz) \\
    \Delta x &= -A^{-1} M (z + \tC_2 v),
    \end{align}\end{subequations}
    where $\tC_2 = A^{-T} C_2$.
    If we condense the block elimination procedure to a single step,
    we obtain
    \begin{equation}
    \Delta x = -A^{-1} M (z - 
        \tC_2 (\tC_2^T M \tC_2)^{-1}(\tC_2^T Ax - d_2 + \tC_2^T Mz)).
    \end{equation}

@@

To deal with the concavity of $f(x)$ we can use a truncated Newton method,
    \begin{equation}
    \hat{\mathcal{L}}(x,p,y) = \mathcal{L}(x,p,y) + \frac{\eta}{2} \|x - x_0\|^2
    \end{equation}
    or even
    \begin{equation}
    \Delta x = -(\nabla_{xx}^2 \mathcal{L} + \eta I)^{-1}
        \nabla_x \mathcal{L},
    \end{equation}
    where the coefficient $\eta$ is increased in order to 
    ensure that the augmented Hessian is positive definite 
    (so that $(\nabla_x\mathcal{L})^T\Delta x > 0$).
However, this generally makes inverting the Hessian intractable.

@@

\subsubsection{Linearly constrained $f(x)$ with arbitrary phase}
We now investigate another special case of $f(x)$, 
    namely, when $f(x)$ implements the equality constraint 
    $C_2^T x = e^{i \theta}d_2$,
    where $C_2 \in \reals^{n \times p_2}$ and $p_2 \ll n$.
This situation corresponds to
    the addition of an arbitrary phase $e^{i\theta}$
    to a linear equality constraint.

We simply note that this case can be handled by computing
    \begin{equation}
    \Delta x = -A^{-1} M (z - 
        \tC_2 (\tC_2^T M \tC_2)^{-1}(\tC_2^T Ax - e^{i\theta}d_2 + \tC_2^T Mz)).
    \end{equation}
    for different choices of $\theta$.
The total computational cost is then simply $n_\theta + p_2$ solves,
    where $n_\theta$ is the number of values of $\theta$ which are tried.

\subsection{Multi-mode ADMM}
Since ADMM explicitly separates updating the field and structure variables,
    handling multiple fields is very naturally included in the formulation.
Namely, multiple fields ($x_i$) are updated simultaneously, in parallel,
    and then are ``connected'' by updating a single $p$ 
    which takes all the $x_i$ into account.

\begin{appendix}
\section{Constructing the relevant matrices and vectors}
The physics residual is based on the electromagnetic wave equation, 
    \begin{equation}
    \nabla \times \epsilon^{-1} \nabla \times H - \mu \omega^2 H - 
        \nabla \times \epsilon^{-1} J - i \omega M = 0,
    \end{equation}
    which can be written as
    \begin{equation}
    r(x,p) = A(p) x - b(x) = B(x) p - d(x).
    \end{equation}

\subsection{Defining the curl operators}
Specifically, the curl operations are defined as
    \begin{equation}
    \nabla^\pm \times u = \begin{bmatrix}
        D_y^\pm u_z - D_z^\pm u_y \\
        D_z^\pm u_x - D_x^\pm u_z \\
        D_x^\pm u_y - D_y^\pm u_x
        \end{bmatrix}
    \end{equation}
    where
    \begin{subequations}\begin{align}
    D_x^- v &= \frac{v(i,j,k) - v(i-1,j,k)}{\phi_x(i-\frac{1}{2})} \\
    D_x^+ v &= \frac{v(i+1,j,k) - v(i,j,k)}{\phi_x(i+\frac{1}{2})},
    \end{align}\end{subequations}
    and likewise for the numerical derivatives in the $y$ and $z$ directions.
Note that we assume that $\Delta x, \Delta y, \Delta z = 1$, and so omit them
    from the denominators of $D$.

Lastly, the $\phi$ functions are used to construct 
    absorbing boundary conditions.
Specifically, we use stretched-coordinate perfectly matched layers\cite{scpml}
    for which $\phi$ is defined as
    \begin{equation}
    \phi_x(x) = \begin{cases}
        1 + \frac{i\sigma}{\omega} x^m & x < t_\text{PML}, \\ 
        1 + \frac{i\sigma}{\omega} (x_\text{max} -x)^m
            & x > x_\text{max} - t_\text{PML}, \\ 
        1 & \text{otherwise,}
        \end{cases}
    \end{equation}
    where $\sigma$ is the strength of the absorbing layer, 
    and $m$ controls the profile of the layer.
Typical values for $\sigma$ and $m$ are $1$ and $2.5$ respectively.

\subsection{Forming the field subproblem}
$A(p)$ and $b(p)$ are constructed as follows,
    \begin{subequations}\begin{align}
    A(p) &= A_1 \diag(\epsilon(p)^{-1}) A_2 - \mu \omega^2 I \\
    b(p) &= A_1 \diag(\epsilon(p)^{-1}) b_j + i \omega b_m \\
    A_1 &= \nabla^- \times \\
    A_2 &= \nabla^+ \times \\
    b_j &= J \\
    b_m &= M,
    \end{align}\end{subequations}
    where $\nabla^+$ and $\nabla^-$ refer to forward- and backward-differencing
    respectively.

\subsection{Forming the structure subproblem}
We now show how to form $r(x,p) = B(x)p - d(x)$.

\section{Solving the matrices}
\section{Linear algebra definitions}
\subsection{Gradient and Hessian calculations}
Mention something about complex values.
\subsection{Matrix inversion lemma}\label{inv lemma}
\begin{equation}
(A + UCV)^{-1} = A^{-1} - A^{-1}U (C^{-1} + V A^{-1} U)^{-1} V A^{-1}
\end{equation}
\subsection{Block elimination of a matrix}\label{block elim}
\begin{equation}
\begin{bmatrix} A_{11} & A_{12} \\ A_{21} & A_{22} \end{bmatrix} 
\begin{bmatrix} x_1 \\ x_2 \end{bmatrix} = 
\begin{bmatrix} b_1 \\ b_2 \end{bmatrix}
\end{equation}
\begin{subequations}\begin{align}
A_{11} x_1 + A_{12} x_2 &= b_1 \\
x_1 &= A_{11}^{-1} (b_1 - A_{12} x_2)
\end{align}\end{subequations}
\begin{subequations}\begin{align}
A_{21} x_1 + A_{22} x_2 &= b_2 \\
A_{21} A_{11}^{-1} (b_1 - A_{12} x_2) + A_{22} x_2 &= b_2 \\
(A_{22} - A_{21} A_{11}^{-1} A_{12}) x_2 &= b_2 - A_{21} A_{11}^{-1} b_1
\end{align}\end{subequations}
Typically, one computes $A_{11}^{-1} A_{12}$ and $A_{11}^{-1} b_1$, 
    and then solves for the rest of the system.
\end{appendix}
\begin{thebibliography}{99}
\bibitem{ADMM} Boyd group ADMM paper
\bibitem{scpml} \url{http://math.mit.edu/~stevenj/18.369/pml.pdf}
\end{thebibliography}
\end{document}
